% ****** Note template ****** %

%\RequirePackage{lineno} 

\documentclass[aps,prd,superscriptaddress,floatfix]{revtex4}


\usepackage{graphicx}  % needed for 
\usepackage{dcolumn}   % needed for some tables
\usepackage{bm}        % for math
\usepackage{amssymb}   % for math

%\def\linenumberfont{\normalfont\small\sffamily}

\usepackage{psfrag}


\def\pp{$p\bar{p}$}


\topmargin=-1.1cm

\begin{document}

%\pacs{}


\title{  
\vspace{0.5cm}
Description of the Fermilab software school data objects
}

\author {Matthew Herndon}

\address{University of Wisconsin, Madison, Wisconsin, Fermi
  National Accelerator Laboratory, Illinois}

%\date{\today}

\begin{abstract}
\vskip 0.5cm
\noindent
This document describes the data objects used in the Fermilab software school.
The objects include a StripSet with raw digitized data for the strip sensors
of the toy simulated detector as well as Hits, Tracks, GenTracks and associated
HitSets, TracksSets and GenTrackSets.  In addition, the formats used to store the
data objects to file are described.
\end{abstract}
\maketitle

%\tableofcontents
%\setpagewiselinenumbers
%%%\modulolinenumbers[10]
%%%\linenumbers

\vspace{0.3cm}

\section{Introduction}
Each objects is described below including the object's class and the methods and
format used to write the objects.  Sets are described with the associated data
object.
\\

The detector geometry of the strip sensors is defined in ~\cite{detectorGeometry}.


\section{The StripSet}
\subsection{StripSet class}
Implemented in the StripSet.hh and StripSet.cc files in the DataObjects directories.
\\

The StripSet contains a std::vector of typedef LayerStripMap with one map per layer of the detector.
The detector is dynamically defined from a geometry file and the vector and size are initialized
at run time.  Each layerStripMap is a map of key int stripNumber and adc value.  This allows
the strip data to be stored in a sparse and ordered format.
\\

Access methods:
\begin{itemize}
\item int numberLayers() To assist in iterating over the vector or directly over the layerMaps
\item strips(): const reference access to the vector using a generic method. The underlying vector container could be 
changed to several different container types without effecting the user interface.
\item layerStrips(int layer): const reference access to a the strip
  data from a single layer. Note that a map is retrieved since maps
  have access methods that are not common to all containers.
\item stripNumber(const iterator) or stripNumber(map pair) to get the
  strip Number
\item stripAdc(const iterator) or stripAdc(map pair) to get the
  strip ADC
\item insertStrip(unsigned int layer, int stripNumber, int adc):  To insert data into the set.
\item print(ostream\&): To print to an a user specified ostream.
\end{itemize}

\subsection{StripSet IO}
Implemented in the StripSetIO.hh and StripSetIO.cc files in the Algorithms directories.
\\

Input and output to disk in controlled by the StripSetIO helper class.  Reading and writing
are performed by the readEvent and writeEvent functions.
\\

The data is stored, or ``streamed'', to disk in bit packed format to illustrate a typical way date is stored in compact
format in many experiments.  The data is stored in individual bytes or pairs of bytes for information
that does not fit in one byte.  The data structure is as follows.
\begin{itemize}
\item Strips: In clear text to identify the Set
\item  1 byte, version: The version is checked on input to make sure the correct streamer code is being used.
\item - 1 byte, layer number: 0-9 (repeated once per layer)
\item - 1 byte, number of strips: max 256
\item -- 2 bytes. strip number max 4048  (these items repeated number of strips times)
\item -- 1 byte. adc max 64 
\end{itemize}

Note layer numbers are not necessary to the
structure but are included to facilitate synchronization with write and read operation..
\\

A number of helper functions exist in the StripHitFunctions file in Geometry directories.  These functions
can convert from strip number and layer to local and global positions and back again and identify whether a strip number is valid
on a given layer.

\subsection{binary input and output}
Data input and output is controlled via standard C++ library binary input and
output functions. std::ofstream, ifstream.  Byte writing is performed for write and read functions as:

\begin{itemize}
\item stripdata.write reinterpret\_cast const char *(\&myInt), 1);
\item stripdata.read (myCharByte, 1);
\end{itemize}
\

\section{The GenHit and GenHitSet}
\subsection{GenHit class}
Implemented in the GenHit.hh and GenHit.cc files in the DataObjects directories.
\\

The GenHit class contains:\\

\begin{itemize}
\item TVector3 \_position
\item int \_layer
\item Int \_trackNumber
\end{itemize}


This class serves for generator GenHits only.\\
\\  

Constructor:

\begin{itemize}
\item Hit(hitPosition,layer,trackNumber)
\end{itemize}

Access methods:
\begin{itemize}
\item position():... names correspond to the data members
\item layer()
\item trackNumber()
\item print(ostream\&): To print to an a user specified ostream.
\end{itemize}

Accessor are named for class variables.  For example position()
retrieves the hit \_position. 

A number of helper functions exist in the StripHitFunctions file in Geometry directories.  These functions
can convert from hit position to local and strip number position or back identify whether a hit position at a sensor
plane is within the active area of the sensor.


\subsection{GenHitSet class}
Implemented in the GenHitSet.hh and GenHitSet.cc files in the DataObjects directories.
\\

The GenHitSet contains a typedef GenHitSetContainer of type std::vector of GenHits.
\\

Access methods:

\begin{itemize}
\item genHits(): const reference access to the vector using a generic method.
The vector index indexes the GenHit number to allow direct access to a hit of a known number.

\item insertGenHit(GenHit): Provided to insert data into the set.

\item print(ostream\&): To print to an a user specified ostream that iterates over
the GenHits calling the print method of each GenHit..
\end{itemize}

\subsection{GenHitSet IO}
Implemented in the GenHitSetIO.hh and GenHitSetIO.cc files in the Algorithms directories.
\\

Input and output to disk in controlled by the GenHitSetIO helper class.  Reading and writing
are performed by the readEvent and writeEvent functions.
\\

The data is stored in text format for simplicity.

\begin{itemize}
\item  Hits: text string to identify the set
\item version: The version is checked on input to make sure the correct streamer code is being used.
\item number of hits
\item hit number (repeated number of hits times)
\item - x position
\item - y position
\item - z position
\item - layer number
\item - track number 
\end{itemize}

\section{The Hit and HitSet}
\subsection{Hit class}
Implemented in the Hit.hh and Hit.cc files in the DataObjects directories.
\\

The Hit class contains:\\

\begin{itemize}
\item TVector3 \_position
\item int \_layer
\item int \_numberStrips
\item int \_charge
\item bool \_goodHit
\item double \_resolution 
\end{itemize}


The class server for reconstructed Hits only.   Associated information such as Hit resolutions
are set using information from DetectorGeometry though a resolution variable exists per hit to allow for any resolution model
to be developed.  Also note that a hit is classified of as ``bad'' if it has a large ADC
value or more than two associated strips indicating that it is actually due to several overlapping hits from
different tracks.
\\

Constructors:

\begin{itemize}
\item Hit(hitPosition,layer,numberStrips,charge,goodhit,resolution)
\end{itemize}

Access methods:
\begin{itemize}
\item TVector3 position():... names correspond to the data members
\item layer()
\item numberStrips()
\item charge()
\item goodHit()
\item resolution
\item print(ostream\&): To print to an a user specified ostream.
\end{itemize}

A number of helper functions exist in the StripHitFunctions file in Geometry directories.  These functions
can convert from hit position to local and strip number position or back identify whether a hit position at a sensor
plane is within the active area of the sensor.
\\

\subsection{HitSet class}
Implemented in the HitSet.hh and HitSet.cc files in the DataObjects directories.
\\

The HitSet contains a typedef HitSetContainer of type std::vector of Hits.
\\

Access methods:

\begin{itemize}
\item hits(): const reference access to the vector using a generic method.
The vector index indexes the Hit number to allow direct access to a hit of a known number.

\item insertHit(Hit): Provided to insert data into the set.

\item print(ostream\&): To print to an a user specified ostream that iterates over
the Hits calling the print method of each Hit..
\end{itemize}

\subsection{HitSet IO}
Implemented in the HitSetIO.hh and HitSetIO.cc files in the Algorithms directories.
\\

Input and output to disk in controlled by the HitSetIO helper class.  Reading and writing
are performed by the readEvent and writeEvent functions.
\\

The data is stored in text format for simplicity.

\begin{itemize}
\item  Hits: text string to identify the set
\item version: The version is checked on input to make sure the correct streamer code is being used.
\item number of hits
\item hit number (repeated number of hits times)
\item - x position
\item - y position
\item - z position
\item - layer number
\item - number of strips
\item - charge
\item - goodHit
\item - resolution
\end{itemize}

\section{The GenTrack and GenTrackSet}
\subsection{GenTrack class}
Implemented in the GenTrack.hh and GenTrack.cc files in the DataObjects directories.
\\

The GenTrack class contains:
\begin{itemize}
\item TLorentzVector \_lorentzVector
\item int \_charge
\item TVector3 \_dr, point of closest approach to the reference point, 0,0,0
\end{itemize}

Constructor:

\begin{itemize}
\item GenTrack(TLorentzVector,int charge, TVector3 dr position of closest approach)
\item{itemize}
\end{itemize}


Access:
\begin{itemize}
\item lorentzVector()::... names correspond to the data members
\item charge()
\item dR(): \_dr

\item makeHelix(TVector3 bField,curvatureC): function which takes the magnetic field and the curvature constant and returns
a Helix which can be used for operations like finding intersections with layers.  Note the magnetic field
value is needed since it may not be uniform and the radius of curvature would be different at different points.
With the standard bField from the DetectorGeometry you get the radius of curvature at the origin.

\item print(ostream\&) to print to an a user specified ostream.
\end{itemize}


\subsection{GenTrackSet class}
Implemented in the GenTrackSet.hh and GenTrackSet.cc files in the DataObjects directories.
\\

The GenTrackSet contains a typedef GenTrackSetContainer of type std::vector of GenTracks.
\\
Access:

\begin{itemize}
\item genTracks(): const reference access to the vector using a generic method.
The vector index indexes the GenTrack number to allow direct access to a hit of a known number.

\item insertTrack(GenTrack): Provided to insert data into the set.

\item print(ostream\&): to print to an a user specified ostream that iterates over
the GenTracks calling the print method of each GenTrack..
\end{itemize}


\subsection{GenTrackSet IO}
Implemented in the GenTrackSetIO.hh and GenTrackSetIO.cc files in the Algorithms directories.
\\

Input and output to disk in controlled by the GenTrackSetIO helper class.  Reading and writing
are performed by the readEvent and writeEvent functions.
\\

The data is stored in text format for simplicity.

\begin{itemize}
\item  GenTracks: text string to identify the set
\item version: The version is checked on input to make sure the correct streamer code is being used.
\item number of GenTracks
\item - GenTrack number (repeated number of GenTrackNumber times)
\item - charge
\item - px
\item - py
\item - pz
\item - E
\item - x position
\item - y position
\item - z position
\end{itemize}

\section{The Track and TrackSet}
\subsection{Track class}
Implemented in the Track.hh and Track.cc files in the DataObjects directories.
\\

The Track class contains:
\begin{itemize}
\item Helix \_helix, 5 track parameter helix
\item int \_covMatrix, 5x5 parameter covariance matrix
\item double \_chi2
\item int \_nDof
\item \_trackHits, a typedef TrackHitContainer of type vector of ints with hit index numbers.
\end{itemize}


Constructor:
\begin{itemize}
\item Tracks are constructed by the full set of parameters above.  These parameters
are generated by the BuildTrack helper function in the Algorithms directories.
That class fits a track based the input hits.
\item note that the Helix class includes the public setHelix() expert method for use in TrackFit where the helix must be updated frequently
\end{itemize}


Access:
\begin{itemize}
\item helix() get Helix object with track parameters\\
The Helix helix parameter follow the convention of~\cite{helix} and are described in detail there.\\
Note that helix parameters are always referenced from (0,0,0) in the parametrization used in this code.
\item TVectorD helixParam(), access to the helix parameter vector
\item Helix parameter accessor functions called  from the helix as
  helix().dR() or directly as dR() ...:
\begin{itemize}
\item dR(): Impact parameter, distance in the x-y place to the reference point (0,0,0) at the point of closest
approach. Signed positive to indicate the reference point is within the circle or negative is outside.
\item phi0(): Phi angle from the reference point to the point of closest approach.
\item kappa(): inverse Pt signed to indicate handedness + right handed (negative charge) - left handed
(positive charge)
\item dZ(): z distance to reference point at point of closest approach
\item tanL(): pZ/pT or cot(theta) where theta is the dip angle which is 90 for track perpendicular
to the z axis and 0 degrees for parallel to the position z axis.
\end{itemize}
\item Other Helix accessors available from helix()
\begin{itemize}
\item alpha() get inverse curvature constant in the magnetic field at the origin
\item radiusOfCurvatureAtOrigin()
\item radiusOfCurvature(TVector3 bField) at an arbitrary point with field bField
\item pT()
\item pZ()
\item cotTheta() same as tanL
\item cosTheta() where theta is the dip angle
\item sinTheta() where theta is the dip angle 
\end{itemize}
\item covMatrix()
\item sigmDr() ... uncertainties for the five track parameters from the covariance matrix
\item charge()
\item chi2()
\item nDof()
\item chi2Prob()
\item lorentzVector() get a root TLorentzVector object based on the track parameters
\item trackHits() get TrackHitContainer vector of hit indices's
\item numberHits()
\item print(ostream\&): to print to an a user specified ostream.

\end{itemize}


\subsection{TrackSet class}
Implemented in the TrackSet.hh and TrackSet.cc files in the DataObjects directories.
\\

The TrackSet is a typedef TrackSetContainer of type std::vector of Tracks.
\\

Access:
\begin{itemize}
\item tracks(): const reference access to the vector using a generic method.
The vector index indexes the Track number to allow direct access to a hit of a known number.

\item insertTrack(Track): provided to insert data into the set.


\item print(ostream\&): to print to an a user specified ostream that iterates over
the Tracks calling the print method of each Track..
\end{itemize}


\subsection{TrackSet IO}
Implemented in the TrackSetIO.hh and TrackSetIO.cc files in the Algorithms directories.
\\

Input and output to disk in controlled by the TrackSetIO helper class.  Reading and writing
are performed by the readEvent and writeEvent functions.
\\

The data is stored in text format for simplicity.

\begin{itemize}
\item Tracks: text string to identify the set
\item version: The version is checked on input to make sure the correct streamer code is being used.
\item number of tracks
\item - Track number (repeated number of trackNumber times)
\item - charge
\item - px
\item - py
\item - pz
\item - E
\item - x position
\item - y position
\item - z position
\item - number of Hits
\item -- Hit index numbers (repeated number of Hits times)
\end{itemize}

The hit list could be used alone to restore the track using BuildTrack.  Currently TrackSetIO is not used.

\begin{thebibliography}{99}

\bibitem{detectorGeometry}
detectorGeometry.pdf note, M. Herndon (2014)

\bibitem{helix} 
  B.~Li, K.~Fujii and Y.~Gao,
  %``Kalman-filter-based track fitting in non-uniform magnetic field with segment-wise helical track model,''
  Comput.\ Phys.\ Commun.\  {\bf 185}, 754 (2014)
  [arXiv:1305.7300 [physics.ins-det]].
  %%CITATION = ARXIV:1305.7300;%%
  %1 citations counted in INSPIRE as of 02 Aug 2014


\end{thebibliography}
% 
\end{document}